---
title: "Case Study 02"
author: "Jeremy Dawkins"
date: "11/19/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 
```

#We will load libraries that we will need in order to use certain functions.
```{r load libraries, include=FALSE}
library(dplyr)  #Allows us to pipe data frames to wrangle data we want to analyze
library(ggplot2) #Allows us to connect data frames with graphs
library(ggpubr) #Allows us to arrange any plots on 1 chart to display
library(hrbrthemes) #Allows us to use the theme function when graphing
library(GGally)    #Allows us to use our correlation matrix
library(Hmisc)    #Allow us to find the correlation between variables with pvalues included
library(olsrr)   #To do a variable/ parameter selection of the predictors
library(caret)   #Use to for the K-fold cross validation of our model
library(car)    #To look at the added variable plots of the regression model
library(e1071) #use of the classification model knn and knn.cv
library(class) #Use of the cross validations techniques
library(Metrics) #Use for the Rmse() function
library(leaps) #use for best predictors variables for salary
```


## Hello,

## As part of DDSAnalytics that specializes in talent management solutions we were hired to look at employee data at a firm to determine factors in attrition at the firm. We will examine various aspects of this dataset using explanatory data set to discover what we can find that's backed up by data. 
#We will conduct an analysis on various information and discover which factors into an employee attrition. 
```{r bring in dataset, echo=FALSE}
#Here I bring in the employee dataset and create a data frame 
Dataset = read.csv("CaseStudy2-data.csv",header = TRUE, sep = ",")

#Let's examine the first 5 rows and the last 5 rows
head(Dataset,5)
tail(Dataset,5)

#Let's examine the structure of the data frame and see if there any missing employee information
str(Dataset)
summary(Dataset)

##We see there are no missing information in the data frame, which is great now we can being our EDA
```


## As we do our EDA let's look first look at the departments
#We will look at the summary to look at many people are in each department. 
#We will go ahead and make factors to use later when trying to graphically represent this data, but also show observations on separate groups.
```{r EDA, echo=FALSE}
#Create factors for the different departments so that we could easily identify how many there are.
Dataset$Department = factor(Dataset$Department)

#Displays how many people are in each department at this particular company
summary(Dataset$Department)

#Create factors for Gender to distinguish how many there are
Dataset$Gender = factor(Dataset$Gender)

#Count how many people in each Job Role
summary(Dataset$JobRole)

#Create factors for Job Role, that we will use later to possibly narrow down which jobs roles are volatile
Dataset$JobRole = factor(Dataset$JobRole)

#Create factors for Marital Status, this could be potentially looked at with people with marriage assuming families
Dataset$MaritalStatus = factor(Dataset$MaritalStatus)

#Create factors for Education Field, examine where the majority people tend to have backgrounds in
Dataset$EducationField = factor(Dataset$EducationField)

#Create factors for Overtime, examine if a person is an exempt or non-exempt employee
Dataset$OverTime = factor(Dataset$OverTime)

#Create factors for Attrition, to compare current and non-current employees
Dataset$Attrition = factor(Dataset$Attrition)
```

#We will examine the data by Departments to focus on any specific detailed information
#We will look at the Sales department specifically here in this section
## We can see here that Attrition did occur in the Sales Department, so we will look at their survey results.
#We discover from the Relationship Plot with Attrition; we discover that people with the Manager Job Role on average did not have well relationships with their superiors. 

#We also wanted to look at the people who currently still works at the firm and how their surveys looked. 
#We discover on average their surveys whether it was relationship satisfaction, job involvement, or job satisfaction they had an overall wonderful experience. 

#So, we would need to keep in the back of our mind, what would cause these employees to leave when the Sales Department as a group tends to have overall high surveys across all Job Levels and Job Roles they have within the department.
#Could this be a money, distance from work, stock options, etc.? We will have to dig further, but for now we will look at this same survey for other departments.

```{r EDA Sales Department, echo=FALSE}
#Let's select only the rows that allows us to look into the Sales Department and if attrition occurred
Dataset_Sales_Y = Dataset %>% filter(Department == "Sales" & Attrition == "Yes")

#Let's look at an employee's response to a survey for job satisfaction and job involvement 
#Nothing here as well, *1 = Not involved , 2 = Somewhat involved, 3 = Involved, 4 = Always Involved
JobInvolvementPlot = Dataset_Sales_Y %>% ggplot(aes(fill = JobRole, y = JobInvolvement, x = JobLevel)) + geom_bar(position = "dodge", stat = "identity") + labs(title = "Job Involvement per Job Role", subtitle = " (Higher Level Equates To More Duties Per Job Role")

#Noting here that *1 = Not Happy, 2-3 = Moderately Happy , 4 = Happy for job satisfaction
JobSatisfactionPlot = Dataset_Sales_Y %>% ggplot(aes(fill = JobRole, y = JobSatisfaction,x = JobLevel)) +geom_bar(position = "dodge", stat = "identity") + labs(title = "Job Satisfaction per Job Role", subtitle = "Job Level (Higher Level Equates To More Duties Per Job Role")

#Here we are looking at the Relationship Satisfaction with who an employee reports to, so their manager or executive. This is important when working at any company because this could lead to issues.
#Here we assume the same ratings where, *1 = Not Happy, 2 = Moderately Happy, 3 = Happy, 4 = Enjoys the Relationship
RelationshipPlot = Dataset_Sales_Y %>% ggplot(aes(x = RelationshipSatisfaction, group = JobRole, fill = JobRole)) + geom_density(adjust = 1.5) + theme_ipsum() + facet_wrap(~JobRole) + labs(title = "Relationship Satisfaction with Manager/Executive", subtitle = "Job Level (Higher Level Equates To More Duties Per Job Role")

#Arrange all three graphs to look at them at the same time. 
ggarrange(JobInvolvementPlot,JobSatisfactionPlot,RelationshipPlot, common.legend = TRUE, ncol = 2, nrow = 2)



#Selecting only the rows that allows us to look into the Sales Department and no attrition occurred
Dataset_Sales_N = Dataset %>% filter(Department == "Sales" & Attrition == "No")

#Let's look at an employee's response to a survey for job satisfaction and job involvement 
#Nothing here as well, *1 = Not involved , 2 = Somewhat involved, 3 = Involved, 4 = Always Involved
JobInvolvementPlot2 = Dataset_Sales_N %>% ggplot(aes(fill = JobRole, y = JobInvolvement, x = JobLevel)) + geom_bar(position = "dodge", stat = "identity") + labs(title = "Job Involvement per Job Role", subtitle = "Job Level (Higher Level Equates To More Responsibilites Per Job Role)")

#Noting here that *1 = Not Happy, 2-3 = Moderately Happy , 4 = Happy for job satisfaction
JobSatisfactionPlot2 = Dataset_Sales_N %>% ggplot(aes(fill = JobRole, y = JobSatisfaction,x = JobLevel)) +geom_bar(position = "dodge", stat = "identity") + labs(title = "Job Satisfaction per Job Role", subtitle = "Job Level (Higher Level Equates To More Responsibilites Per Job Role)")

#Here we are looking at the Relationship Satisfaction with who an employee reports to, so their manager or executive. This is important when working at any company because this could lead to issues.
#Here we assume the same ratings where, *1 = Not Happy, 2 = Moderately Happy, 3 = Happy, 4 = Enjoys the Relationship
RelationshipPlot2 = Dataset_Sales_N %>% ggplot(aes(y = RelationshipSatisfaction, x = JobLevel, fill = JobRole)) + geom_violin() + labs(title = "Relationship Satisfaction with Manager/Executive", subtitle = "Job Level (Higher Level Equate To More Responsibilites Per Job Role)")

#Arrange all three graphs to look at them at the same time. 
ggarrange(JobInvolvementPlot2,JobSatisfactionPlot2,RelationshipPlot2, common.legend = TRUE, ncol = 2, nrow = 2)
```


#In this section we will examine the Research & Development Department. We would want to look at people who did leave the company if any, and the people who are currently with them.
#The Job Involvement for former employees on average their surveys seem to be between 2-3 (Moderately Happy) on average regardless of the job levels for each job role. This could be a potential reason factor in employees leaving, we would have to examine further on Job Involvement.
#We also look at the Job Satisfaction where the job levels 2, 3 and 5 seems to have low survey scores. Job Satisfaction overall could lead to a factor on why people would leave because they would feel unhappy with their responsibilities. 
#The relationship Satisfaction between employee and manager or executive seems to over lab across all job levels with the average leaning below 3 (Moderately happy). We see that Research Scientist specifically is overall job level 1 seems to have scored two.
```{r EDA R&D Department, echo = FALSE}
#We will filter by department (R&D), so we can only look at R&D and former employees
Dataset_RD_Y = Dataset %>% filter(Department == "Research & Development" & Attrition == "Yes")

#Let's look at an employee's response to a survey for job satisfaction and job involvement 
#Nothing here as well, *1 = Not involved , 2 = Somewhat involved, 3 = Involved, 4 = Always Involve
JobInvolvementPlot3 = Dataset_RD_Y %>% ggplot(aes(fill = JobRole, y = JobInvolvement, x = JobLevel)) + geom_bar(position = "dodge", stat = "identity") + labs(title = "Job Involvement per Job Role", subtitle = "Job Level (Higher Level Equates To More Responsibilites Per Job Role")

#Noting here that *1 = Not Happy, 2-3 = Moderately Happy , 4 = Happy for job satisfaction
JobSatisfactionPlot3 = Dataset_RD_Y %>% ggplot(aes(fill = JobRole, y = JobSatisfaction, x = JobLevel)) + geom_bar(position = "dodge", stat = "identity") + labs(title = "Job Satisfaction per Job Role", subtitle = "Job Level (Higher Level Equates To More Responsibilites Per Job Role)")


#Here we are looking at the Relationship Satisfaction with who an employee reports to, so their manager or executive. This is important when working at any company because this could lead to issues.
#Here we assume the same ratings where, *1 = Not Happy, 2 = Moderately Happy, 3 = Happy, 4 = Enjoys the Relationship
RelationshipPlot3 = Dataset_RD_Y %>% ggplot(aes(y = RelationshipSatisfaction, x = JobLevel, fill = JobRole)) + geom_violin() + labs(title = "Relationship Satisfaction with Manager/Executive", subtitle = "Job Level (Higher Level Equate To More Responsibilites Per Job Role)")


#Arrange all three graphs to look at them at the same time. 
ggarrange(JobInvolvementPlot3,JobSatisfactionPlot3,RelationshipPlot3, common.legend = TRUE, ncol = 2, nrow = 2)


#We will filter by deparment (R&D) and current employees
Dataset_RD_N = Dataset %>% filter(Department == "Research & Development" & Attrition == "No")

#Let's look at an employee's response to a survey for job satisfaction and job involvement 
#Nothing here as well, *1 = Not involved , 2 = Somewhat involved, 3 = Involved, 4 = Always Involve
JobInvolvementPlot4 = Dataset_RD_N %>% ggplot(aes(fill = JobRole, y = JobInvolvement, x = JobLevel)) + geom_bar(position = "dodge", stat = "identity") + labs(title = "Job Involvement per Job Role", subtitle = "Job Level (Higher Level Equates To More Responsibilites Per Job Role")


#Noting here that *1 = Not Happy, 2-3 = Moderately Happy , 4 = Happy for job satisfaction
JobSatisfactionPlot4 = Dataset_RD_N %>% ggplot(aes(fill = JobRole, y = JobSatisfaction, x = JobLevel)) + geom_bar(position = "dodge", stat = "identity") + labs(title = "Job Satisfaction per Job Role", subtitle = "Job Level (Higher Level Equates To More Responsibilites Per Job Role)")


#Here we are looking at the Relationship Satisfaction with who an employee reports to, so their manager or executive. This is important when working at any company because this could lead to issues.
#Here we assume the same ratings where, *1 = Not Happy, 2 = Moderately Happy, 3 = Happy, 4 = Enjoys the Relationship
RelationshipPlot4 = Dataset_RD_N %>% ggplot(aes(y = RelationshipSatisfaction, x = JobLevel, fill = JobRole)) + geom_violin() + labs(title = "Relationship Satisfaction with Manager/Executive", subtitle = "Job Level (Higher Level Equate To More Responsibilites Per Job Role)")


#Arrange all three graphs to look at them at the same time. 
ggarrange(JobInvolvementPlot4,JobSatisfactionPlot4,RelationshipPlot4, common.legend = TRUE, ncol = 2, nrow = 2)

```


#In this section we will examine the HR Department. We would want to look at similar aspect compared to the other two departments we examined. 
#In this section we see there are only two job roles in this department, one labeled Human Resources and the other labeled Manager
#We also see the people who are now former employees are only Human Resources Job Roles. 
#Their overall Relationship Satisfaction between employee and managers on average seems to be around the grade 3 (Happy). We will need to examine further what other factors caused these employees to leave.
```{r EDA HR Department, echo = FALSE}
#We will filter by department (HR), so we can only want to look at HR former employees
Dataset_HR_Y = Dataset %>% filter(Department == "Human Resources" & Attrition == "Yes")

#Let's look at an employee's response to a survey for job satisfaction and job involvement 
#Nothing here as well, *1 = Not involved , 2 = Somewhat involved, 3 = Involved, 4 = Always Involve
JobInvolvementPlot5 = Dataset_HR_Y %>% ggplot(aes(fill = JobRole, y = JobInvolvement, x = JobLevel)) + geom_bar(position = "dodge", stat = "identity") + labs(title = "Job Satisfaction per Job Role", subtitle = "Job Level (Higher Level Equates To More Responsibilites Per Job Role")


#Noting here that *1 = Not Happy, 2-3 = Moderately Happy , 4 = Happy for job satisfaction
JobSatisfactionPlot5 = Dataset_HR_Y %>% ggplot(aes(fill = JobRole, y = JobSatisfaction, x = JobLevel)) + geom_bar(position = "dodge", stat = "identity") + labs(title = "Job Satisfaction per Job Role", subtitle = "Job Level (Higher Level Equates To More Responsibilites Per Job Role)")


#Here we are looking at the Relationship Satisfaction with who an employee reports to, so their manager or executive. This is important when working at any company because this could lead to issues.
#Here we assume the same ratings where, *1 = Not Happy, 2 = Moderately Happy, 3 = Happy, 4 = Enjoys the Relationship
RelationshipPlot5 = Dataset_HR_Y %>% ggplot(aes(y = RelationshipSatisfaction, x = JobLevel, fill = JobRole)) + geom_violin() + labs(title = "Relationship Satisfaction with Manager/Executive", subtitle = "Job Level (Higher Level Equate To More Responsibilites Per Job Role)")


#Arrange all three graphs to look at them at the same time. 
ggarrange(JobInvolvementPlot5,JobSatisfactionPlot5,RelationshipPlot5, common.legend = TRUE, ncol = 2, nrow = 2)



#We will filter by department (HR), so we can only look at HR current employees
Dataset_HR_N = Dataset %>% filter(Department == "Human Resources" & Attrition == "No")

#Let's look at an employee's response to a survey for job satisfaction and job involvement 
#Nothing here as well, *1 = Not involved , 2 = Somewhat involved, 3 = Involved, 4 = Always Involve
JobInvolvementPlot6 = Dataset_HR_N %>% ggplot(aes(fill = JobRole, y = JobInvolvement, x = JobLevel)) + geom_bar(position = "dodge", stat = "identity") + labs(title = "Job Satisfaction per Job Role", subtitle = "Job Level (Higher Level Equates To More Responsibilites Per Job Role")


#Noting here that *1 = Not Happy, 2-3 = Moderately Happy , 4 = Happy for job satisfaction
JobSatisfactionPlot6 = Dataset_HR_N %>% ggplot(aes(fill = JobRole, y = JobSatisfaction, x = JobLevel)) + geom_bar(position = "dodge", stat = "identity") + labs(title = "Job Satisfaction per Job Role", subtitle = "Job Level (Higher Level Equates To More Responsibilites Per Job Role)")


#Here we are looking at the Relationship Satisfaction with who an employee reports to, so their manager or executive. This is important when working at any company because this could lead to issues.
#Here we assume the same ratings where, *1 = Not Happy, 2 = Moderately Happy, 3 = Happy, 4 = Enjoys the Relationship
RelationshipPlot6 = Dataset_HR_N %>% ggplot(aes(y = RelationshipSatisfaction, x = JobLevel, fill = JobRole)) + geom_violin() + labs(title = "Relationship Satisfaction with Manager/Executive", subtitle = "Job Level (Higher Level Equate To More Responsibilites Per Job Role)")


#Arrange all three graphs to look at them at the same time. 
ggarrange(JobInvolvementPlot6,JobSatisfactionPlot6,RelationshipPlot6, common.legend = TRUE, ncol = 2, nrow = 2)
```


#After discovering interesting facts about variables that could be factors in attrition. Now we will look at majority of the variables here to see which variables are correlated between each other. If these variables are correlated between each other then we would need to take them out and not use them in our potential model. This is since multiple factors trying to explain our model in the next section will not achieve anything and could be overfitting the model. We were able to find a model that best fit and look at the evidence to back up our model. We will also conduct a classification model with Naive Bayes to see sensitivity and specificity. 
```{r Correlation Factors, echo = FALSE}
#We will take subset the initial Dataset we brought
Dataset_Y = Dataset 

#We will not select numerical variables from this data frame to use in a correlation matrix. 
#Let's look at the column names to use in our correlation matrix
colnames(Dataset_Y)

#We will take our characters in the Business Travel category, so we can make these numeric to include it in our correlation matrix
Dataset_Y2 = gsub("_","",Dataset_Y$BusinessTravel)
Dataset_Y$BusinessTravel = gsub("-","",Dataset_Y2)

#Make Business Travel and Attrition a factor here. 
Dataset_Y$BusinessTravel = factor(Dataset_Y$BusinessTravel)
Dataset_Y$Attrition = factor(Dataset_Y$Attrition)


#We will convert the Marital Status of an employee to a numeric value, where *3 = Single, 2 = Married, 1 = Divorced
#We will convert the Gender Status of an employee to a numeric value, where *1 = Female, 2 = Male
#Business Travel is converted to an numeric, where *1 = No Travel , 2 = Travel Frequently , 3 = Travel Rarely
#Department is also converted to an numeric, where *1 = Human Resources, 2 = Research & Development, 3 = Sales
#Over Time is converted to an numeric, where *1 = No , 2 = Yes
Dataset_Y = Dataset_Y %>% mutate(MaritalStatus2 = as.numeric(MaritalStatus), Gender2 = as.numeric(Gender), BusinessTravel2 = as.numeric(BusinessTravel), Department2 = as.numeric(Department), OverTIme2 = as.numeric(OverTime), Attrition2 = as.numeric(Attrition))


#Let's select columns that we want to add to our correlation matrix, and leave out the non numeric variables
Dataset_Y_Select = Dataset_Y %>% select(Attrition2,Age,DailyRate,DistanceFromHome,Education,EnvironmentSatisfaction,HourlyRate,JobInvolvement,JobLevel,JobSatisfaction,MonthlyIncome,MonthlyRate,PercentSalaryHike,PerformanceRating,RelationshipSatisfaction,StockOptionLevel,TotalWorkingYears,TrainingTimesLastYear,WorkLifeBalance,YearsAtCompany,YearsInCurrentRole,YearsSinceLastPromotion,YearsWithCurrManager,MaritalStatus2,Gender2,BusinessTravel2,Department2,OverTIme2,Department)

#Let's check to make sure we have all the numeric variables we want to consider as factors for Attrition
#Get the summary statistics as well
str(Dataset_Y_Select)
summary(Dataset_Y_Select)


#Let's create a visualization of correlation with all the variable we believe are factors in attrition
Corr_Matrix = ggcorr(Dataset_Y_Select[,c(1:28)], method = c("everything","pearson")) 


#We will create a model that best represents our scatter plot matrix, which in this case would be all the predictors in the Multiple Linear Regression Model
fit = lm(Attrition2~.-Department, data = Dataset_Y_Select)
summary(fit)

#we will now look at three different selection predictor process, we will compare each one to choose the best model, in which we will later use for model testing.

#First selection method is forward selection, we will use a pvalue of 0.05 at the cut off point. 
FS = ols_step_forward_p(fit,penter = 0.05,details = TRUE)


#Our second method is backward selection, which we will use the same pvalue of 0.05
BS = ols_step_backward_p(fit, prem = 0.05,details = TRUE)


#Our last method is the stepwise selection method with pvalues of 0.05
SW = ols_step_both_p(fit,pent = 0.05,prem = 0.05,details = TRUE)


#Now we will compare to select which model is best using different Selection methods
#We will look at the R square, Adjusted R2, Overall F Test, RSME, MSE,MAE,AIC, SBC
FS$rsquare[11]
BS$rsquare[14]
SW$rsquare[11]

FS$sbc[11]
BS$sbc[14]
SW$sbc[11]

FS$adjr[11]
BS$adjr[14]
SW$adjr[11]

FS$rmse[11]
BS$rmse[14]
SW$rmse[11]

FS$aic[11]
BS$aic[14]
SW$aic[11]

FS$mallows_cp[11]
BS$mallows_cp[14]
SW$mallows_cp[11]

#After comparing the model it appears the backward selection process will be used based on the smaller errors. 
#We will conduct the fit of the model and check assumptions for MLR
fit2 = lm(Attrition2~DistanceFromHome+EnvironmentSatisfaction+JobInvolvement+JobLevel+JobSatisfaction+TrainingTimesLastYear+WorkLifeBalance+YearsInCurrentRole+YearsSinceLastPromotion+MaritalStatus2+Department2+OverTIme2, data = Dataset_Y_Select)
summary(fit2)

#We will evaluate residual plots and assumptions to of this model
#We will assume the assumptions for regression is met. 
par(mfrow=c(1,2))
hist(fit2$residuals, col = "blue", main = "Histogram of Residuals", xlab = "Residuals")
plot(fit2$fitted.values, fit2$residuals, main = "Plot of Residuals vs Fitted Values", xlab = "Fitted Values", ylab = "Residuals")

#We will conduct a K-fold Cross Validation to evaluate the model performance of our data. 
#We will split the data into training and a test set. We will create a training set containing 75 percent of the data and a test set containing the other 25 percent. 
set.seed(123)
training_obs = sample(seq(1,dim(Dataset_Y_Select)[1]),round(.75 * dim(Dataset_Y_Select)[1]), replace = FALSE)
Train_Dataset = Dataset_Y_Select[training_obs,]

Test_Dataset = Dataset_Y_Select[-training_obs,]


#Here we are using our model that we made to use on the training data set. We will also look at the summary details of this model. 
model1_fit = lm(Attrition2~DistanceFromHome+EnvironmentSatisfaction+JobInvolvement+JobLevel+JobSatisfaction+TrainingTimesLastYear+WorkLifeBalance+YearsInCurrentRole+YearsSinceLastPromotion+MaritalStatus2+Department2+OverTIme2, data = Train_Dataset)
summary(model1_fit)

#We will use the prediction function using our model and test on the testing dataset
model1_pred = predict(model1_fit,newdata = Test_Dataset)

#We will create a data frame and that will house the Rsquared, RSME, MAE
data.frame(Rsq = R2(model1_pred, Test_Dataset$Attrition2), RSME = RMSE(model1_pred, Test_Dataset$Attrition2), MAE = MAE(model1_pred, Test_Dataset$Attrition2))

#Display the prediction error rate, which we would want small as possible
RMSE(model1_pred, Test_Dataset$Attrition2) / mean(Test_Dataset$Attrition2)


#We will read in the initial set again
Initial_Set = read.csv("CaseStudy2-data.csv",header = TRUE, sep = ",")

#Create a new data frame object using the initial set
DF = Initial_Set
DF = Initial_Set %>% select(Age,BusinessTravel,DailyRate,Department,DistanceFromHome,Education,EducationField,EnvironmentSatisfaction,Gender,HourlyRate,JobInvolvement,JobLevel,JobRole,JobSatisfaction,MaritalStatus,MonthlyIncome,MonthlyRate,NumCompaniesWorked,Over18,OverTime,PercentSalaryHike,PerformanceRating,RelationshipSatisfaction,TotalWorkingYears,TrainingTimesLastYear,WorkLifeBalance,YearsAtCompany,YearsInCurrentRole,YearsSinceLastPromotion,YearsWithCurrManager,Attrition)


#We will not use Naive Bayes to conduct a classification algorithm to see if a employee is attrited or not based on multiple predictors. 
#We would want to split our data into a training and a test set to train the data then test our results. 
#We will try to make the split as random as possible. 
set.seed(1234)
Index = sample(seq(1:dim(DF)[1]),round(.75 * dim(DF)[1]), replace = FALSE)
Class_SetTrain = DF[Index,]
Class_SetTest = DF[-Index,]

#Here we create our naive bayes model with our training set using the predictors we determined and see if it can correctly match what we have for attrition
#Once that is done we will create a table and a confusion matrix to see how accurate how classification is, as well as the sensitivity and specificity
model = naiveBayes(Class_SetTrain[,c(1:30)], Class_SetTrain$Attrition)
help = table(predict(model,Class_SetTest[,c(1:30)]), Class_SetTest$Attrition)
confusionMatrix((help))


#We will subset our initial data set from the csv file we pulled, and only focus on specific factors for our regression model 
Dataset_Sal = subset(Initial_Set, select = c(2,5,7:8,11:12,14:16,18,20:22,25:27,29:36)) 
Dataset_Sal = Dataset_Sal %>% mutate(Salary = MonthlyIncome * 12) #Find the yearly Salary for all employees 

linear_model = lm(Salary~.-MonthlyIncome,data = Dataset_Sal) #use all factors as explanatory variables for the regression model

best_subset = regsubsets(Salary~.-MonthlyIncome, data = Dataset_Sal, nbest = 1, nvmax = NULL, force.in = NULL, force.out = NULL, method = "exhaustive") #uses subset to figure out the best explanatory variables to use in the model

summary_best_subset <- summary(best_subset) #Finds the summary of the subset of by marking which variables to use with "True" or "False"

which.max(summary_best_subset$adjr2) #finds the highest adjusted r square
summary_best_subset$which[8,] #finds the predictors with the 8 


#Cross Validation of the Salary Model we made
#We will use our linear model and set a partition of 50% for the train and 50% of the test data. 
set.seed(1234)
training.samples <- Dataset_Sal$Salary %>% createDataPartition(p = 0.5, list = FALSE)
train.data  <- Dataset_Sal[training.samples, ]
test.data <- Dataset_Sal[-training.samples, ]
# We build the linear model that best described Salary
model <- lm(Salary~DistanceFromHome+EnvironmentSatisfaction+JobInvolvement+JobLevel+PercentSalaryHike+PerformanceRating+TotalWorkingYears+YearsWithCurrManager, data = Dataset_Sal)
# Make predictions and compute the R2, RMSE and MAE and displays it as a data frame
predictions <- model %>% predict(test.data)
data.frame( R2 = R2(predictions,test.data$Salary),
            RMSE = RMSE(predictions, test.data$Salary),
            MAE = MAE(predictions, test.data$Salary))

#Calculates the prediction error rate
RMSE(predictions, test.data$Salary)/mean(test.data$Salary)


#Another Cross- Validation called "Leave one out cross validation" 
#We will create a train control with the specific cross validation method
train.control <- trainControl(method = "LOOCV")
#Train the model that we have selected
model2 <- train(Salary ~DistanceFromHome+EnvironmentSatisfaction+JobInvolvement+JobLevel+PercentSalaryHike+PerformanceRating+TotalWorkingYears+YearsWithCurrManager, data = Dataset_Sal, method = "lm",
               trControl = train.control)
#Summarize the results of the cross validation
print(model2)

```


#We will not take the factors that we believe are big impacts to attrition and create different models with them
#We will show visuals of the multiple linear regression for Attrition and Salary
```{r Modelings with Factors and plots, echo = FALSE}
#We will look at the Individual plots of the explanatory variable and the response for our regression model. 
avPlots(fit2)

#We will look at the individual plots of the explanatory variables and the response for our regression model with Salary. 
avPlots(linear_model)

#Here is a density chart of the various Salarys within each Job Level
ggplot(Dataset_Sal, aes(x=JobLevel, y=Salary) ) +
  stat_density_2d(aes(fill = ..level..), geom = "polygon", colour="white") + labs(title = "Salary vs Job Level Density Chart")

#Here is a chart looking at performance rating and salary
ggplot(Dataset_Sal, aes(x=JobInvolvement, y=Salary) ) +
  stat_density_2d(aes(fill = ..level..), geom = "polygon", colour="green") + labs(title = "Salary vs Job Involvement Density Chart")


```


#We will demonstrate some graphs of our models that we created. We display any visuals that the employer can see on what are some low factors they can take into account. 
```{r Summary (with Graphs), echo = FALSE}
#Here are a few charts that model some areas the employer can try to improve at. 
proportion = table(Dataset_Y$Attrition) / nrow(Dataset_Y)
#Here we display a boxplot of Job Satisfaction and Attrition
boxplot(Dataset_Y$JobSatisfaction ~ Dataset_Y$Attrition ,width = proportion,
    ylab="Job Satisfaction" , xlab="- Attrition -", col = c("orange","seagreen"), main = "Boxplot of Job Satisfaction of Attrition")



#We will write our prediction model to a csv and push it to github. 
write.csv(model1_pred,file = "prediction.attrition.csv")
write.csv(predictions, file = "prediction.salary.csv")
```


#In conclusion, we discovered that the main factors that leads to attrition are Job Satisfaction, Job Level, Job Involvement, OverTime, Martial Status, and Years in Current Role. These appear to have great significance in whether a employee would stay with the company or leave. Perhaps the company can focus on these areas of weakness and discover what is cause low survey scores that is causing attrition. There could be number of reasons, it will be difficult to narrow it down to only one. The former employees could be poor performers in the workplace which could have an impact on productivity. Any company would not want high attrition rate due to it's cost like hiring replacements and training them. 